Supplementary Methods for\\ \emph{Effectiveness of UNAIDS targets and HIV vaccination across 127 countries}

### 2016-11-19 Documents/supplementary_information.tex
A few tweaks.

\renewcommand{\thesection}{S\arabic{section}}
\renewcommand{\figurename}{Supplementary Fig.}
\renewcommand{\figureautorefname}{\figurename}
\renewcommand{\tablename}{Supplementary Table}
\renewcommand{\tableautorefname}{\tablename}

@@

\author{Jan Medlock}
\affil{Department of Biomedical Sciences, Oregon State University, 106
  Dryden Hall, Corvallis, OR, 97331-4801, USA,
  \href{mailto:jan.medlock@oregonstate.edu}{\texttt{jan.medlock@oregonstate.edu}}}
\author{Abhishek Pandey} \author{Alyssa S.~Parpia} \author{Amber Tang}
\author{Laura A.~Skrip} \author{Alison P.~Galvani}
\affil{Center for Infectious Disease Modeling and Analysis, Yale
  School of Public Health, 135 College Street, New Haven, USA}

@@

We developed a country-level continuous-time, compartmental model that
divides people aged $15$\;y and older into non-overlapping HIV states.
We chose to consider only these ages because they are the vast
majority of global PLHIV \autocite{UNICEF}, 93\% in 2014; because
population-level HIV estimates and demographic data are often
stratified into ages $15$--$49$; and to capture the continuing burden
of infections in people ages $50$ and older.

The background mortality rate ($\mu$) in each of the
127 countries was collected from published estimates
\autocite{World_Development_Indicators2013-ee}.  The recruitment rate
($\kappa$) into the age $15\;\text{y}$ and older group was set by the
difference between the country's population growth rate
\autocite{WorldBankpg} and its background mortality rate
\autocite{World_Development_Indicators2013-ee}, as would be true if the
population was at its stable age structure.

Our model has 8 HIV states
(\hyperref[model_diag]{Extended Data Figs.~\ref*{model_diag}}):
susceptible to HIV infection ($S$ is the number of people
susceptible), vaccinated against HIV ($R$), acute HIV infection ($A$),
undiagnosed HIV infection ($U$), diagnosed but untreated HIV infection
($D$), treated without viral suppression ($T$), viral suppression
($V$), and having AIDS ($W$).  Viral suppression is defined as a viral
load of less than 1000 viral RNA copies per mL of blood.  Viral
suppression not only extends survival but also dramatically reduces
transmission
\autocite{vernazza2000, ioannidis2001, attia2009, May2014-gp}.
Transitions between model states are governed by a series of
differential equations parametrized using values found in other
studies, along with transmission rates derived from UNAIDS estimates
of incidence and prevalence (\autoref{model_param}).  The model is
deterministic, but uncertainty in the parameters is treated by running
the model 1000 times with samples from the parameter distributions
(\autoref{uncertainty}) and summarizing the model outcomes using the
median, 1st and 3rd quartiles (i.e., 25th and 75th percentiles), and
5th and 95th percentiles.  Each country was modeled separately
(\hyperref[effectiveness_Afghanistan]{Extended Data
  Figs.~\ref*{effectiveness_Afghanistan}}--\ref{effectiveness_Zimbabwe})
and the country simulations were aggregated to calculate global and
regional outcomes
(\hyperref[effectiveness_Global]{Extended Data
  Figs.~\ref*{effectiveness_Global}}--\ref{effectiveness_Western_Europe}).

\begin{table}
  \begin{center}
    \begin{tabularx}{\textwidth}{lXlll}
      \hline
      & Definition & Value & Reference \\
      \hline
      $\mu$ & Background death rate
      & Country specific, Supplementary Table 2
      & \autocite{World_Development_Indicators2013-ee} \\
      $\kappa$ & Recruitment rate
      & Country specific, Supplementary Table 2
      & \autocite{World_Development_Indicators2013-ee, WorldBankpg} \\
      $\delta$	& Rate of leaving acute infection
      & $\Triangular$(2, 4.14, 9.6)\;y$^{-1}$
      & \autocite{Hollingsworth2008-iy} \\
      $\sigma$	& Rate of developing AIDS without viral suppression
      & 0.1064\;y$^{-1}$ & \autocite{Morgan2002-cq} \\
      $\theta$ & Rate of developing AIDS with viral suppression
      & Eq.~(\ref{theta}) & --- \\
      $\gamma$ & Rate of viral suppression
      & $\Uniform$(0.5, 1.5)\;y$^{-1}$
      & \autocite{Currie2009-yz} \\
      $\nu$ & Death rate from AIDS & 0.5\;y$^{-1}$
      & \autocite{Morgan2002-cq} \\
      $\rho$ & Vaccine efficacy & 50\%\,[30\%, 70\%] & --- \\
      $\omega$	& Reduction in lifetime with viral suppression
      & $\Uniform$(5, 8)\;y
      & \autocite{Samji2013-kf, Unaids2014-ue} \\
      $\tau_{A}$ & Transmissibility during acute phase
      & $\Triangular$(0.0039, 0.0082, 0.0150)
      & \autocite{Wawer2005-us, Skarbinski2015-ni}\\
      $\tau_{U}$ & Transmissibility after acute phase
      & $\Triangular$(0.00077, 0.0014, 0.00251)
      & \autocite{Hughes2012-so} \\
      $\varepsilon$ & Relative transmissibility with
      viral suppression & $\BetaPERT$(0.08, 0.002, 0.57)
      & \autocite{Donnell2010-xo, attia_2009, wilson_2012, jia_2013,
        rodger_2016} \\
      $n$ & Coital acts per year & $\Uniform$(96, 108)
      & \autocite{Wawer2005-us, Abdool_Karim2010-cm}\\
      $\alpha$ & Diagnosis rate & See eq.~(\ref{diagnosis_rate}) & --- \\
      $\phi$ & Treatment rate & See eq.~(\ref{treatment_rate}) & --- \\
      $\psi$ & Rate of relapse to untreated & See eq.~(\ref{relapse_rate})
      & --- \\
      $\zeta$ & Vaccination rate & See eq.~(\ref{vaccination_rate}) & --- \\
      \hline
    \end{tabularx}
    \caption{Model parameters. $\Triangular$, $\Uniform$, and
      $\BetaPERT$ are sampling distributions: see
      \autoref{uncertainty} for a full description.  For vaccine
      efficacy, $50\%$ was the baseline, and $30\%$ and $70\%$ were
      also used in a scenario analysis.}
    \label{model_param}
  \end{center}
\end{table}

People susceptible to HIV ($S$) progress to acute infection ($A$)
according to the force of infection ($\lambda$,
eq.~\eqref{force_of_infection}), which depends on the HIV transmission
rate, estimated from the country-specific incidence and prevalence
data (\autoref{model_fitting}), adjusted for the current number of
acutely infected, untreated or ineffectively treated, and viral
suppression PLHIV.  Acute infection is characterized by high viral
loads and lasts an average of 2.90 [95\% confidence interval 1.23,
6.00]\;months \autocite{Hollingsworth2008-iy}.  (We converted these
durations to rates like $1 / 2.90\;\text{month$^{-1}$} \approx
4.14\;\text{y$^{-1}$}$ and set the parameter using random samples from
$\Triangular(2, 4.14, 9.6)\;\text{y$^{-1}$}$ to capture the
uncertainty in its value.  See \autoref{uncertainty} for the
definition of $\Triangular(a, b, c)$ and other distributions.)  We
assumed that people remain undiagnosed during their acute-infection
phase. After acute infection, people move into the chronic undiagnosed
class ($U$).  People in the undiagnosed class will be diagnosed at the
rate $\alpha$, which is determined by the country's current diagnosis
level and its diagnosis target (\autoref{targets}).  Diagnosed people
($D$) transition to the treated compartment ($T$) upon starting ART at
the rate $\phi$, dependent on the country's current treatment level
and its treatment target (\autoref{targets}).  After some time,
treated people then transition to the viral-suppression class ($V$),
at the rate $\gamma$.  We took $\gamma$ as a random sample from
$\Uniform(0.5, 1.5)\;\text{y$^{-1}$}$, that is, viral suppression
occurs after between 8 months and 2 years on treatment
\autocite{Currie2009-yz}.  Disengagement from treatment moves people from
treated ($T$) and viral suppression ($V$) at the rate $\psi$,
dependent on the current level of viral suppression and the country's
target (\autoref{targets}), back to the diagnosed class ($D$).

PLHIV without viral suppression (compartments $U$, $D$, and $T$)
develop AIDS (compartment $W$) after an average duration of 10.4 years
\autocite{Morgan2002-cq}, i.e., with transition rate
$\sigma = 0.1064\;\text{y$^{-1}$}$.  People who have achieved viral
suppression ($V$) develop AIDS at rate $\theta$, slower than PLHIV
without viral suppression.  People with viral suppression are
estimated to live $5$--$8$\;y shorter lives than the non-HIV infected
\autocite{Samji2013-kf, Unaids2014-ue}, which we quantified by
randomly sampling the parameter $\omega$ from
$\Uniform(5, 8)\;\text{y}$.  Assuming no relapse to untreated, the
duration of viral suppression is
\begin{equation}
  \frac{1}{\theta + \mu} = \frac{1}{\mu} - \omega - \frac{1}{\nu}.
\end{equation}
The term to the left of the equals sign is the time until leaving
viral suppression by either progression to AIDS (at rate $\theta$) or
background mortality (at rate $\mu$).  The term to the right of the
equals sign is the lifespan of the non-HIV infected, minus the
reduction in life due to having HIV with viral suppression, minus the
duration of the AIDS stage.  Solving for $\theta$ gives the rate of
progression to AIDS with viral suppression as
\begin{equation}
  \label{theta}
  \theta = \frac{1}{1/\mu - \omega - 1/\nu} - \mu.
\end{equation}

@@

The model equations are
\begin{equation}
  \label{model_eqns}
  \begin{split}
    \frac{\md S}{\md t} &= \kappa N - \lambda S - \zeta S- \mu S,
    \\
    \frac{\md R}{\md t} & = \zeta S - (1 - \rho) \lambda R - \mu R,
    \\
    \frac{\md A}{\md t} &= \lambda S + (1 - \rho) \lambda R - \delta A - \mu A,
    \\
    \frac{\md U}{\md t} &= \delta A - \alpha U - \mu U - \sigma U,
    \\
    \frac{\md D}{\md t} &=  \alpha U + \psi T + \psi V
    - \phi D - \mu D - \sigma D,
    \\
    \frac{\md T}{\md t} &= \phi D - \psi T - \gamma T - \mu T
    - \sigma T,
    \\
    \frac{\md V}{\md t} &= \gamma T - \psi V - \mu V - \theta V,
    \\
    \frac{\md W}{\md t} &= \sigma U + \sigma D + \sigma T + \theta V -
    \nu W,
    \\
    \frac{\md Y}{\md t} &= \lambda \big[S + (1 - \rho) R\big],
    \\
    \frac{\md Z}{\md t} &= \nu W,
  \end{split}
\end{equation}
with force of infection
\begin{equation}
  \label{force_of_infection}
  \lambda =
  \frac{\eta \big[\beta_A A + \beta_U (U + D + T) + \beta_V V\big]}{N},
\end{equation}
sexually active population size $N = S + R + A + U + D + T + V$.  We
assumed that people with AIDS ($W$) are too sick to be sexually
active.

@@

The parametrized differential equations were solved numerically
\autocite[see][]{medlock2016-git} and four effectiveness outcomes were
computed from the solutions.
\begin{description}
\item[New infections] New infections from 2015 to time $t$ are given
  by $Y(t)$.

@@

The proportion of people on ART who have achieved viral suppression is
\begin{equation}
  p_V(t) = \frac{V(t)}{T(t) + V(t)}.
\end{equation}
The rate of people relapsing to untreated is
\begin{equation}
  \label{relapse_rate}
  \psi(t) = \psi_{\max} H\big(p_V(t) - p_V^*(t)\big),
\end{equation}
with $\psi_{\max} = 1$, which, unlike diagnosis and treatment rates,
prevents relapses ($\psi(t) = 0$) when the current level of viral
suppression is below the target ($p_V^*(t) > p_V(t)$) and allows
relapses ($\psi(t) > 0$) when the current level of viral suppression
is at or above target.  We chose to vary the relapse rate to achieve
the desired level of viral suppression because achieving viral
suppression seems to depend on the time on treatment (i.e.,
$1 / \gamma$), while relapse rates are difficult to estimate.

@@

For 127 countries, we found sufficient published estimates and data to
parametrize our model.  Demographic data, including population growth
rate \autocite{WorldBankpg}, death rate
\autocite{World_Development_Indicators2013-ee}, and number of people
aged 15--49 years \autocite{The_World_Bank2016-fd} were obtained from
the World Bank (Supplementary Table 2). Longitudinal HIV prevalence
(Supplementary Table 3) and incidence (Supplementary Table 4)
estimates for ages 15--49, spanning from as early as 1990 to 2015,
were primarily obtained from the AIDSinfo database produced by UNAIDS
\autocite{Unaids2016-an}. Other sources included UNAIDS Country
Progress Reports \autocite{Unaids2016-am} published from 2012 to 2016
and AIDS Data Hub \autocite{AIDSdatahub-fg}, which compiles data from
UNAIDS, UNICEF, WHO, and the Asian Development Bank. These sources
were also used to inform the initial conditions for the model: the
number of people diagnosed with HIV, the number on ART, and the number
who have viral suppression or have been retained on treatment for at
least 12 months (Supplementary Table 5). Where estimates or data were
not available from these sources, alternative resources such as
peer-reviewed journal articles and country health ministry reports
were consulted.  See Supplementary Tables 2--5 for full information on
the sources used for each country.  These source tables are also
available in the public source-code repository
\autocite{medlock2016-git}.

@@

A few countries—e.g., Bulgaria, Timor-Leste, and Yemen—are currently
estimated by UNAIDS to have low HIV prevalence but relatively high
incidence, resulting in large estimated transmission rates, leading to
rapid projected growth of their HIV epidemics by 2035. The apparent
low prevalence and high incidence may, at least in part, be due to the
resolution in UNAIDS estimates (0.1\% for prevalence, 0.01\% for
annual per-capita incidence).  Results for such countries should be
considered in light of this limitation.

@@

Projections were generated for each country through model simulations
using 1000 random samples from published parameter distributions
(\autoref{model_param}) and the country-specific estimated
distribution of the transmission rate (\autoref{model_fitting}). The
parameters were sampled using Latin hypercube sampling
\autocite{blower1994}.  We also used the variance reduction technique of
running different interventions with the same parameter samples, i.e.,
one set of sample parameter values is drawn, status quo and
intervention scenarios are all run with those sample values and
compared, and this process is repeated for each of the 1000 sample
sets of parameters \autocite{shechter2006}.

The model outcomes were summarized using the median, 1st and 3rd
quartiles (i.e., 25th and 75th percentiles), and 5th and 95th
percentiles.  We calculated partial rank correlation coefficients
(PRCC) \autocite{blower1994} to measure the independent effect of
each non-vaccine parameter on model projections of global new HIV
infections from 2015 to 2035
(\hyperref[PRCCs]{Extended Data Fig.~\ref*{PRCCs}}).

@@

\item[Beta-PERT] $\BetaPERT(a, b, c)$ is the Beta-PERT probability
  distribution \autocite{malcom1959} with minimum $a$, mode $b$, and
  maximum $c$.  It has density function
  \begin{equation}
    \label{BetaPERT}
    f_{\BetaPERT}(x) =
    \begin{cases}
      \frac{(x - a)^{v - 1} (c - x)^{w - 1}}{(c - a)^{v + w - 2} B(v, w)}
      & \text{if $a \leq x \leq c$,}
      \\
      0 & \text{otherwise,}
    \end{cases}
  \end{equation}
  where
  \begin{equation}
    \begin{split}
      \mu &= \frac{a + \lambda b + c}{\lambda + 2},
      \\
      v &= \frac{(\mu - a)(2 b - a - c)}{(b - \mu) (c - a)},
      \\
      w &= \frac{v (a - \mu)}{\mu - c},
    \end{split}
  \end{equation}
  and $B(v, w)$ is the standard beta function \autocite[\S 6.2]{davis1972}.
  We used the standard value of $\lambda = 4$.

>>>>

\renewcommand{\thesection}{S\arabic{section}}
\newcommand{\sectionname}{Section}
\renewcommand{\sectionautorefname}{\sectionname}
\renewcommand{\figurename}{Supplementary Fig.}
\renewcommand{\figureautorefname}{\figurename}
\renewcommand{\tablename}{Supplementary Table}
\renewcommand{\tableautorefname}{\tablename}

@@

\author{Jan Medlock}
\affil{Department of Biomedical Sciences, Oregon State University, 106
  Dryden Hall, Corvallis, OR, 97331-4801, USA,
  \href{mailto:jan.medlock@oregonstate.edu}
  {\texttt{jan.medlock@oregonstate.edu}}}
\author{Abhishek Pandey} \author{Alyssa S.~Parpia} \author{Amber Tang}
\author{Laura A.~Skrip} \author{Alison P.~Galvani}
\affil{Center for Infectious Disease Modeling and Analysis, Yale
  School of Public Health, 135 College Street, New Haven, USA}

@@

We developed a country-level continuous-time mathematical model that
divides people aged $15$\;y and older into non-overlapping HIV states.
We focus on this age group because it accounts for the vast majority
of global PLHIV \autocite{UNICEF}, 93\% in 2014; because
population-level HIV estimates and demographic data are typically only
estimated for ages $15$--$49$; and to capture the continuing burden of
infections in people ages $50$ and older.

Our model has 8 HIV states (Extended Data Fig.~\ref*{model_diag}):
susceptible to HIV infection ($S$ is the number of people
susceptible), vaccinated against HIV ($R$), acute HIV infection ($A$),
undiagnosed HIV infection ($U$), diagnosed but untreated HIV infection
($D$), treated without viral suppression ($T$), viral suppression
($V$), and having AIDS ($W$).  Viral suppression is defined as a viral
load of less than 1000 viral RNA copies per mL of blood.  Viral
suppression not only extends survival but also dramatically reduces
transmission
\autocite{vernazza2000, ioannidis2001, attia2009, May2014-gp}.

Transitions between model states are governed by a system of
differential equations parametrized using values found in other
studies, along with transmission rates derived from UNAIDS estimates
of incidence and prevalence (\autoref{model_param}).  The model is
deterministic, but uncertainty in the parameters was treated by
running the model 1000 times with samples from the parameter
distributions (\autoref{uncertainty}) and summarizing the model
outcomes using the median and percentiles.  Each country was modeled
separately (Extended Data
Figs.~\ref*{effectiveness_Afghanistan}--\ref*{effectiveness_Zimbabwe})
and the country simulations were aggregated to calculate global and
regional outcomes (Extended Data
Figs.~\ref*{effectiveness_Global}--\ref*{effectiveness_Western_Europe}).

The background mortality rate ($\mu$) in each of the 127 countries was
collected from published estimates
\autocite{World_Development_Indicators2013-ee}.  The recruitment rate
($\kappa$) into the age $15\;\text{y}$ and older group was set by the
difference between the country's population growth rate
\autocite{WorldBankpg} and its background mortality rate
\autocite{World_Development_Indicators2013-ee}, as would be true if
the population was at its stable age structure.  (More on these data
sources is in \autoref{data_sources}.)

\begin{table}
  \begin{center}
    \begin{tabularx}{\textwidth}{lXlll}
      \hline
      & Definition & Value & Reference \\
      \hline
      $\mu$ & Background death rate
      & Country specific, Supplementary Table 2
      & \autocite{World_Development_Indicators2013-ee} \\
      $\kappa$ & Recruitment rate
      & Country specific, Supplementary Table 2
      & \autocite{World_Development_Indicators2013-ee, WorldBankpg} \\
      $\delta$	& Rate of leaving acute infection
      & $\Triangular$(2, 4.14, 9.6)\;y$^{-1}$
      & \autocite{Hollingsworth2008-iy} \\
      $\sigma$	& Rate of developing AIDS without viral suppression
      & 0.1064\;y$^{-1}$ & \autocite{Morgan2002-cq} \\
      $\theta$ & Rate of developing AIDS with viral suppression
      & Eq.~(\ref{theta}) & --- \\
      $\gamma$ & Rate of viral suppression
      & $\Uniform$(0.5, 1.5)\;y$^{-1}$
      & \autocite{Currie2009-yz} \\
      $\nu$ & Death rate from AIDS & 0.5\;y$^{-1}$
      & \autocite{Morgan2002-cq} \\
      $\rho$ & Vaccine efficacy & 50\%\,[30\%, 70\%] & --- \\
      $\omega$	& Reduction in lifetime with viral suppression
      & $\Uniform$(5, 8)\;y
      & \autocite{Samji2013-kf, Unaids2014-ue} \\
      $\tau_{A}$ & Transmissibility during acute phase
      & $\Triangular$(0.0039, 0.0082, 0.0150)
      & \autocite{Wawer2005-us, Skarbinski2015-ni}\\
      $\tau_{U}$ & Transmissibility after acute phase
      & $\Triangular$(0.00077, 0.0014, 0.00251)
      & \autocite{Hughes2012-so} \\
      $\varepsilon$ & Relative transmissibility with
      viral suppression & $\BetaPERT$(0.08, 0.002, 0.57)
      & \autocite{Donnell2010-xo, attia_2009, wilson_2012, jia_2013,
        rodger_2016} \\
      $n$ & Coital acts per year & $\Uniform$(96, 108)
      & \autocite{Wawer2005-us, Abdool_Karim2010-cm}\\
      $\alpha$ & Diagnosis rate & See eq.~(\ref{diagnosis_rate}) & --- \\
      $\phi$ & Treatment rate & See eq.~(\ref{treatment_rate}) & --- \\
      $\psi$ & Rate of relapse to untreated & See eq.~(\ref{relapse_rate})
      & --- \\
      $\zeta$ & Vaccination rate & See eq.~(\ref{vaccination_rate}) & --- \\
      \hline
      \vspace*{-2.5em}
    \end{tabularx}
    \caption{Model parameters. $\Triangular$, $\Uniform$, and
      $\BetaPERT$ are sampling distributions (\autoref{uncertainty}).
      For vaccine efficacy, $50\%$ was the baseline, and $30\%$ and
      $70\%$ were also used in a scenario analysis.}
    \label{model_param}
  \end{center}
\end{table}

People susceptible to HIV ($S$) progress to acute infection ($A$)
according to the force of infection ($\lambda$), which depends on the
HIV transmission rate, estimated from the country-specific incidence
and prevalence data (\autoref{model_fitting}), adjusted for the
current number of acutely infected, untreated or ineffectively
treated, and viral suppression PLHIV.  Acute infection is
characterized by high viral loads and lasts an average of 2.90 [95\%
confidence interval 1.23, 6.00]\;months
\autocite{Hollingsworth2008-iy}.  (We converted these durations to
rates like $1 / 2.90\;\text{month$^{-1}$} \approx
4.14\;\text{y$^{-1}$}$ and set the parameter using random samples from
$\Triangular(2, 4.14, 9.6)\;\text{y$^{-1}$}$ to capture the
uncertainty in its value.  See \autoref{uncertainty} for the
definition of $\Triangular(a, b, c)$ and other distributions.)  We
assumed that people remain undiagnosed during their acute-infection
phase. After acute infection, people move into the chronic undiagnosed
class ($U$).  People in the undiagnosed class will be diagnosed at the
rate $\alpha$, which is determined by the country's current diagnosis
level and its diagnosis target (\autoref{targets}).  Diagnosed people
($D$) transition to the treated compartment ($T$) upon starting ART at
the rate $\phi$, dependent on the country's current treatment level
and its treatment target (\autoref{targets}).  After some time,
treated people then transition to the viral-suppression class ($V$),
at the rate $\gamma$.  We took $\gamma$ as a random sample from
$\Uniform(0.5, 1.5)\;\text{y$^{-1}$}$, that is, viral suppression
occurs after between 8 months and 2 years on treatment
\autocite{Currie2009-yz}.  Disengagement from treatment moves people
from treated ($T$) and viral suppression ($V$) at the rate $\psi$,
dependent on the current level of viral suppression and the country's
target (\autoref{targets}), back to the diagnosed class ($D$).

PLHIV without viral suppression (compartments $U$, $D$, and $T$)
develop AIDS (compartment $W$) after an average duration of 10.4 years
\autocite{Morgan2002-cq}, i.e. with transition rate
$\sigma = 0.1064\;\text{y$^{-1}$}$.  People who have achieved viral
suppression ($V$) develop AIDS at rate $\theta$, slower than PLHIV
without viral suppression.  People with viral suppression are
estimated to live $5$--$8$\;y shorter lives than the non-HIV infected
\autocite{Samji2013-kf, Unaids2014-ue}, which we quantified by
randomly sampling the parameter $\omega$ from
$\Uniform(5, 8)\;\text{y}$.  Assuming no relapse to untreated, the
duration of viral suppression is
\begin{equation}
  \frac{1}{\theta + \mu} = \frac{1}{\mu} - \omega - \frac{1}{\nu}.
\end{equation}
The term to the left of the equals sign is the time until leaving
viral suppression by either progression to AIDS (at rate $\theta$) or
background mortality (at rate $\mu$).  The term to the right of the
equals sign is the lifespan of the non-HIV infected, minus the
reduction in life due to having HIV with viral suppression, minus the
duration of the AIDS stage.  Solving for $\theta$ gives the rate of
progression to AIDS with viral suppression as
\begin{equation}
  \label{theta}
  \theta = \frac{1}{1/\mu - \omega - 1/\nu} - \mu.
\end{equation}

@@

The model equations are
\begin{equation}
  \label{model_eqns}
  \begin{split}
    \frac{\md S}{\md t} &= \kappa N - \lambda S - \zeta S- \mu S,
    \\
    \frac{\md R}{\md t} & = \zeta S - (1 - \rho) \lambda R - \mu R,
    \\
    \frac{\md A}{\md t} &= \lambda S + (1 - \rho) \lambda R - \delta A - \mu A,
    \\
    \frac{\md U}{\md t} &= \delta A - \alpha U - \mu U - \sigma U,
    \\
    \frac{\md D}{\md t} &=  \alpha U + \psi T + \psi V
    - \phi D - \mu D - \sigma D,
    \\
    \frac{\md T}{\md t} &= \phi D - \psi T - \gamma T - \mu T
    - \sigma T,
    \\
    \frac{\md V}{\md t} &= \gamma T - \psi V - \mu V - \theta V,
    \\
    \frac{\md W}{\md t} &= \sigma U + \sigma D + \sigma T + \theta V -
    \nu W,
    \\
    \frac{\md Y}{\md t} &= \lambda \big[S + (1 - \rho) R\big],
    \\
    \frac{\md Z}{\md t} &= \nu W,
  \end{split}
\end{equation}
with force of infection
\begin{equation}
  \label{force_of_infection}
  \lambda =
  \frac{\eta \big[\beta_A A + \beta_U (U + D + T) + \beta_V V\big]}{N},
\end{equation}
and sexually active population size $N = S + R + A + U + D + T + V$.
We assumed that people with AIDS ($W$) are too sick to be sexually
active.

@@

The parametrized differential equations were solved numerically using
the LSODA routine \autocite{odepack, scipy, medlock2016-git} and four
effectiveness outcomes were computed from the solutions.
\begin{description}
\item[New infections] New infections from 2015 to time $t$ are given
  by $Y(t)$.

@@

The proportion of people on ART who have achieved viral suppression is
\begin{equation}
  p_V(t) = \frac{V(t)}{T(t) + V(t)}.
\end{equation}
The rate of people relapsing to untreated is
\begin{equation}
  \label{relapse_rate}
  \psi(t) = \psi_{\max} H\big(p_V(t) - p_V^*(t)\big),
\end{equation}
with $\psi_{\max} = 1$, which, unlike diagnosis and treatment rates,
prevents relapses ($\psi(t) = 0$) when the current level of viral
suppression is below the target ($p_V^*(t) > p_V(t)$) and allows
relapses ($\psi(t) > 0$) when the current level of viral suppression
is at or above target.  We chose to vary the relapse rate to achieve
the desired level of viral suppression because achieving viral
suppression seems to depend on the time on treatment (i.e.
$1 / \gamma$), while relapse rates are difficult to estimate.

@@

For 127 countries, we found sufficient published estimates and data to
parametrize our model.  Demographic data, including population growth
rate \autocite{WorldBankpg}, death rate
\autocite{World_Development_Indicators2013-ee}, and number of people
aged 15--49 years \autocite{The_World_Bank2016-fd} were obtained from
the World Bank (Supplementary Table 2). Longitudinal HIV prevalence
(Supplementary Table 3) and incidence (Supplementary Table 4)
estimates for ages 15--49, spanning from as early as 1990 for some
countries to 2015, were primarily obtained from the AIDSinfo database
produced by UNAIDS \autocite{Unaids2016-an}. Other sources included
UNAIDS Country Progress Reports \autocite{Unaids2016-am} published
from 2012 to 2016 and AIDS Data Hub \autocite{AIDSdatahub-fg}, which
compiles data from UNAIDS, UNICEF, WHO, and the Asian Development
Bank. These sources were also used to inform the initial conditions
for the model: the number of people diagnosed with HIV, the number on
ART, and the number who have viral suppression or have been retained
on treatment for at least 12 months (Supplementary Table 5). Where
estimates or data were not available from these sources, alternative
resources such as peer-reviewed journal articles and country health
ministry reports were consulted.  See Supplementary Tables 2--5 for
full information on the sources used for each country.  These source
tables are also available in the public source-code repository
\autocite{medlock2016-git}.

@@

A few countries—e.g. Bulgaria, Timor-Leste, and Yemen—are currently
estimated by UNAIDS to have low HIV prevalence but relatively high
incidence, resulting in large estimated transmission rates, leading to
rapid projected growth of their HIV epidemics by 2035. The apparent
low prevalence and high incidence may, at least in part, be due to the
resolution in UNAIDS estimates (0.1\% for prevalence, 0.01\% for
annual per-capita incidence).  Results for such countries should be
considered in light of this limitation.

@@

Projections were generated for each country through model simulations
using 1000 random samples from published parameter distributions
(\autoref{model_param}) and the country-specific estimated
distribution of the transmission rate (\autoref{model_fitting}). The
parameters were sampled using Latin hypercube sampling
\autocite{blower1994}.  We also used the variance reduction technique of
running different interventions with the same parameter samples, i.e.
one set of sample parameter values is drawn, status quo and
intervention scenarios are all run with those sample values and
compared, and this process is repeated for each of the 1000 sample
sets of parameters \autocite{shechter2006}.

The model outcomes were summarized using the median, 1st and 3rd
quartiles (i.e. 25th and 75th percentiles), and 5th and 95th
percentiles.  We calculated partial rank correlation coefficients
(PRCC) \autocite{blower1994} to measure the independent effect of each
non-vaccine parameter on model projections of global new HIV
infections from 2015 to 2035 (Extended Data Fig.~\ref*{PRCCs}).

@@

\item[Beta-PERT] $\BetaPERT(a, b, c)$ is the Beta-PERT probability
  distribution \autocite{malcom1959} with minimum $a$, mode $b$, and
  maximum $c$.  It has density function
  \begin{equation}
    \label{BetaPERT}
    f_{\BetaPERT}(x) =
    \begin{cases}
      \frac{(x - a)^{v - 1} (c - x)^{w - 1}}{(c - a)^{v + w - 2} B(v, w)}
      & \text{if $a \leq x \leq c$,}
      \\
      0 & \text{otherwise,}
    \end{cases}
  \end{equation}
  where
  \begin{equation}
    \begin{split}
      \mu &= \frac{a + \lambda b + c}{\lambda + 2},
      \\
      v &= \frac{(\mu - a)(2 b - a - c)}{(b - \mu) (c - a)},
      \\
      w &= \frac{v (a - \mu)}{\mu - c},
    \end{split}
  \end{equation}
  and $B(v, w)$ is the standard beta function \autocite{davis1972}.
  % \cite[\S6.2]{davis1972}
  We used the standard value of $\lambda = 4$.
